---
title: "In Season Model"
author: "Josh Kroman"
output:
  html_document: default
  theme: flatly
vignette: "%\\VignetteIndexEntry{Stock Recruit Model} %\\VignetteEngine{knitr::rmarkdown}
  %\\VignetteEncoding{UTF-8}\n"
#output: word_document
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 5
)
```

```{r setup, include = FALSE}
library(ggplot2)
library(tidyverse)
library(bayesplot)
library(wesanderson)
library(SRJPEdata)
library(SRJPEmodel)
```

## Overview

The in-season outmigrant abundance model aims to predict the Juvenile Production Estimate (JPE) in near-real time by leveraging historical relationships between partial season catch estimates and total seasonal outmigration. 

Unlike the stock-recruit approach, which relies on adult spawner estimates, the in-season model allows forecasting within the same water year. Note: these forecasts become more reliable later in the season as more data are available.

## Model Objective

To develop a predictive model that estimates total spring-run Chinook outmigrant abundance at Delta entry using partial in-season RST data.

## Conceptual model

The framework fits historical timing curves to weekly BT‑SPAS‑X outputs (including uncertainty), then uses partial-season observations to forecast remaining weekly passage, total outmigrants, and downstream routing to the Delta.

*Figure 1: Conceptual Diagram of SRJPE In - Season model.* 

# Model Architecture

# Model Inputs

# Running stock recruit submodel


# Model Output



**Resources**
