---
title: "Stock Recruit Model"
author: "Josh Kroman"
output:
  html_document: default
  theme: flatly
vignette: "%\\VignetteIndexEntry{Stock Recruit Model} %\\VignetteEngine{knitr::rmarkdown}
  %\\VignetteEncoding{UTF-8}\n"
#output: word_document
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 5
)
```

```{r setup, include = FALSE}
library(ggplot2)
library(tidyverse)
library(bayesplot)
library(wesanderson)
library(SRJPEdata)
library(SRJPEmodel)
```

## Overview

Freshwater stock-recruitment models with covariates will be used to estimate annual outmigrant abundance as a function of spawner abundance and covariates. The models will be fit to historical estimates of abundance and covariate conditions. These relationships will then be used to make forecasts of outmigrant abundance in the JPE model. 

## Model Objective

## Conceptual model



*Figure 1: Conceptual Diagram of SRJPE model. Survival sub-model shown in yellow.* 

# Model Architecture 

Spawner abundances used in the fitting can be direct annual observations of total redds (from redd surveys) or spawners (from holding or carcass surveys), or from models that predict spawner abundance based on passage abundance (the P2S model). Water year type, discharge, or water temperature will be used as covariates in the modeling. Covariates consist of annual values that will capture effects of environmental conditions on survival rates for different components of early life history between spawning and outmigration (e.g., water temperature during incubation, emergence, or early rearing). Owing to limitations in sample size (number of years available) and uncertainty in stock (spawner) and recruitment (outmigrant abundance) estimates, we anticipate estimating single-covariate models only. An information theoretic approach will be used to identify the most predictive models. We will develop tributary-specific stock-recruitment models as well as models for the mainstem traps based on the sum of spawner abundances in tributaries upstream of the mainstem traps.

# Model Inputs

# Running stock recruit submodel


# Model Output



**Resources**

